### 2022-05-12 notebooks/compare-racmo.Rmd
Adds notebook comparing SMB estimates in wAIS for RACMO and PAIPR

---
title: "Comparison of WAIS accumulation trends in PAIPR to RACMO"
output: 
  html_notebook:
    code_folding: hide
---

```{r}
library(here)
library(dplyr)
library(tidyr)
library(broom)
library(ggplot2)
library(sf)
library(cetcolor)
```

```{r}
dat = readRDS(here('data/Rdata-clean.rds'))
dat.gdf = readRDS(here('data/Rdata-gdf_trace.rds'))

# Remove rows with no RACMO SMB or PAIPR accum estimates
dat = dat %>% drop_na(SMB, accum)

# Remove high accumulation outliers
dat = dat %>% filter(accum <= 1000)

# Only select traces with full coverage 1979-2009
yr.cutoff = 2008
dat = dat %>% group_by(trace_ID) %>% filter(Year >= 1979) %>% filter(Year <= yr.cutoff) %>% 
  filter(all(1979:yr.cutoff %in% Year))

# Remove (now) missing traces from gdf
dat.gdf = dat.gdf %>% filter(trace_ID %in% dat$trace_ID)

# Add mean SMB results to gdf
dat.gdf = left_join(dat.gdf %>% select(-accum), 
                    dat %>% group_by(trace_ID) %>% summarise(accum=mean(accum), SMB=mean(SMB)))
dat.gdf = dat.gdf %>% mutate(accum.res = SMB-accum)

# Calculate residuals
dat = dat %>% mutate(res = SMB-accum)
```

Direct 1:1 plot of PAIPR to RACMO (averaged over all years).
```{r}
ggplot(dat.gdf) + geom_point(aes(x=accum, y=SMB), color='blue', alpha=0.25) + 
  geom_smooth(aes(x=accum, y=SMB), method = 'lm') + geom_abline(slope=1, color='red') + 
  lims(x=c(100,600), y=c(100,600))

ggplot(dat.gdf) + geom_sf(aes(color=accum.res)) + scale_color_gradient2()
```

```{r PAIPR-RACMO_plots}
# ggplot(dat) + geom_density(aes(x=res, group=Year, fill=Year), color=NA, alpha=0.2) + 
#   scale_fill_viridis_c() + 
#   geom_density(aes(x=res), color='black')

ggplot(dat, aes(x=factor(Year), y=res)) + 
  geom_hline(yintercept = 0) + 
  geom_violin(aes(fill=Year), position="dodge", alpha=0.5) + 
  stat_summary(fun=median, geom="point", size=1, color="red") + 
  scale_fill_viridis_c()
```

```{r}
accum.glm = dat %>% 
  mutate(Year = Year-mean(Year)) %>% 
  group_by(trace_ID) %>% 
  do(tidy(glm(accum ~ Year, family=Gamma(link='log'), data = .))) %>% 
  filter(term=='Year') %>% select(-term) %>% mutate(est.trans = exp(estimate)-1)
SMB.glm = dat %>% 
  mutate(Year = Year-mean(Year)) %>% 
  group_by(trace_ID) %>% 
  do(tidy(glm(SMB ~ Year, family=Gamma(link='log'), data = .))) %>% 
  filter(term=='Year') %>% select(-term) %>% mutate(est.trans = exp(estimate)-1)

dat.gdf = inner_join(dat.gdf, 
                     accum.glm %>% select(trace_ID, est.trans) %>% rename(estimate=est.trans), 
                     by="trace_ID", suffix=c(".raw", ".accum")) %>% 
  left_join(SMB.glm %>% select(trace_ID, est.trans) %>% rename(estimate=est.trans), 
            by="trace_ID", suffix=c(".accum", ".SMB"))
```

```{r}
ggplot(dat.gdf) + geom_density(aes(x=estimate.accum), fill='blue', alpha=0.5) + 
  geom_density(aes(x=estimate.SMB), fill='red', alpha=0.5)
```

```{r}
dat.gdf = dat.gdf %>% mutate(trend.res = estimate.accum-estimate.SMB)
ggplot(dat.gdf) + geom_density(aes(x=trend.res), fill='blue', alpha=0.5) + 
  geom_vline(xintercept = 0)

ggplot(dat.gdf) + 
  geom_sf(aes(color=trend.res))  + 
  # scale_color_gradientn(colours = cet_pal(7, name = "d1"))
  scale_color_gradient2()
```

```{r}
ggplot(dat %>% group_by(Year) %>% summarise(SMB=median(SMB), accum=median(accum))) + 
  geom_line(aes(x=Year, y=accum), color='blue') + 
  geom_line(aes(x=Year, y=SMB), color='red')

ggplot(dat %>% group_by(trace_ID)) + 
  geom_line(aes(x=Year, y=accum, group=trace_ID), color='blue', alpha=0.01) + 
  geom_line(aes(x=Year, y=SMB, group=trace_ID), color='red', alpha=0.01)
```

```{r}
ggplot(dat.gdf) + geom_sf(aes(color=trace_ID)) + scale_color_viridis_c()
```

```{r}
ri = dat$trace_ID[sample.int(nrow(dat),1)]
ggplot(dat %>% filter(trace_ID==ri)) + 
  geom_line(aes(x=Year, y=accum), color='blue') + 
  geom_line(aes(x=Year, y=SMB), color='red') + 
  ggtitle(label = ri)
```