### 2022-05-27 tutorials/sp-regress.Rmd
Adds notebook for testing spacetime modeling examples with Kronecker sums for effects

---
title: "R Notebook"
output: html_notebook
---

This is a reproduction of the example in the [section on dynamic space-time regression](https://becarioprecario.bitbucket.io/spde-gitbook/ch-stapp.html#dynamic-space-time-regression) in "Advanced Spatial Modeling with Stochastic Partial Differential Equations Using R and INLA".
I do this as an understanding check prior to adapting and applying it to my data.

```{r}
library(INLA)
library(dplyr)
library(tidyr)
library(broom)
library(ggplot2)
library(spdep)
```

Define custom function (found in `spde-book-files/R/spde-book-functions.R`) for generated spatially correlated data.
```{r}
book.rMatern <- function(n, coords, sigma=1, range, kappa = sqrt(8*nu)/range, variance = sigma^2, nu=1) {
    m <- as.matrix(dist(coords))
    m <- exp((1-nu)*log(2) + nu*log(kappa*m)-
             lgamma(nu))*besselK(m*kappa, nu)
    diag(m) <- 1
    return(drop(crossprod(chol(variance*m),
                          matrix(rnorm(nrow(coords)*n), ncol=n))))
}
```

Generate random spatial locations
```{r}
n <- 150
set.seed(1)
coo <- matrix(runif(2 * n), n)
```

Generate spatially and temporally correlated data
```{r}
kappa <- c(10, 12)
sigma2 <- c(1 / 2, 1 / 4)
k <- 15
rho <- c(0.7, 0.5)

set.seed(2)
beta0 <- book.rMatern(k, coo, range = sqrt(8) / kappa[1],
  sigma = sqrt(sigma2[1]))

set.seed(3)
beta1 <- book.rMatern(k, coo, range = sqrt(8) / kappa[2],
  sigma = sqrt(sigma2[2]))
beta0[, 1] <- beta0[, 1] / (1 - rho[1]^2)
beta1[, 1] <- beta1[, 1] / (1 - rho[2]^2)

for (j in 2:k) {
  beta0[, j] <- beta0[, j - 1] * rho[1] + beta0[, j] *
    (1 - rho[1]^2)
  beta1[, j] <- beta1[, j - 1] * rho[2] + beta1[, j] *
    (1 - rho[2]^2)
}

set.seed(4)
# Simulate the covariate values
hh <- runif(n * k) 
mu.beta <- c(-5, 1)
taue <- 20

set.seed(5)
# Error in the observation
error <- rnorm(n * k, 0, sqrt(1 / taue))

# Dynamic regression part
y <- (mu.beta[1] + beta0) + (mu.beta[2] + beta1) * hh + 
  error

# DGK Model year as covariate
year.val = matrix(data=(1:k)/k, nrow=n, ncol=k, byrow=TRUE)
# y = (mu.beta[1] + beta0) + (mu.beta[2] + beta1) * year.val + 
#   error

# Arrange in tbl
dat = tibble(loc.idx=rep(1:n,k), E=rep(coo[,1],k), N=rep(coo[,2],k), Year=rep(1:k, each=n), y=c(y), hh=hh, yr.val = c(year.val))
```

Exploration of the data.
```{r}
h.trends = dat %>% 
  group_by(loc.idx) %>% 
  do(tidy(lm(y ~ hh, data = .))) %>% 
  filter(term=='hh') %>% select(-term)
dat.mu = dat %>% group_by(loc.idx) %>% 
  summarize(E=mean(E), N=mean(N), y=mean(y)) %>% 
  left_join(h.trends %>% select(loc.idx, estimate))

ggplot(dat.mu) + geom_point(aes(x=E, y=N, color=y)) + scale_color_viridis_c()
ggplot(dat.mu) + geom_point(aes(x=E, y=N, color=estimate)) + scale_color_gradient2()
ggplot(dat %>% group_by(loc.idx)) + geom_line(aes(x=Year, y=y, group=loc.idx), alpha=0.1)
```

```{r}
mesh <- inla.mesh.2d(dat %>% select(E,N), max.edge = c(0.15, 0.3),
  offset = c(0.05, 0.3), cutoff = 0.07)

plot(mesh)
points(dat %>% select(E, N), col = "red")

spde <- inla.spde2.pcmatern(mesh = mesh,
  prior.range = c(0.05, 0.01), # P(practic.range < 0.05) = 0.01
  prior.sigma = c(1, 0.01)) # P(sigma > 1) = 0.01

i0 <- inla.spde.make.index('i0', spde$n.spde, n.group = length(unique(dat$Year)))
i1 <- inla.spde.make.index('i1', spde$n.spde, n.group = length(unique(dat$Year)))
```

```{r}
A0 <- inla.spde.make.A(mesh, loc=cbind(dat$E, dat$N), group = dat$Year)
A1 <- inla.spde.make.A(mesh, loc=cbind(dat$E, dat$N), 
                       group = dat$Year, weights = dat$yr.val)

dat.stack <- inla.stack(
  data = list(y = dat$y), 
  A = list(A0, A1, 1), 
  effects = list(i0, 
                 i1, 
                 tibble(Intercept = 1, mu1 = 1, 
                        yr.val=dat$yr.val, h = dat$hh)),
  tag = 'dat')

form <- y ~ 0 + mu1 + yr.val + # to fit mu_beta
  f(i0, model = spde, group = i0.group, 
    control.group = list(model = 'ar1')) + 
  f(i1, model = spde, group = i1.group, 
    control.group = list(model = 'ar1'))

# Assigned initial starting values for parameter estimates
theta.ini <- c(
  log(taue), # likelihood log precision
  log(sqrt(8) / kappa[1]), # log range 1
  log(sqrt(sigma2[1])), # log stdev 1
  log((1 + rho[1])/(1 - rho[1])), # log trans. rho 1
  log(sqrt(8) / kappa[2]), # log range 1
  log(sqrt(sigma2[2])), # log stdev 1
  log((1 + rho[2]) / (1 - rho[2])))# log trans. rho 2
```

```{r model, cache=TRUE}
mod = inla(form, family = 'gaussian',
           data = inla.stack.data(dat.stack), 
           control.predictor = list(A = inla.stack.A(dat.stack)), 
           control.mode = list(theta = theta.ini, restart=TRUE), # initial theta value
           control.inla = list(int.strategy = 'eb'))# no integr. wrt theta)
```

Extract model results of interest
```{r}
# Get indices of observed data
fitted.idx = inla.stack.index(dat.stack, 'dat')$data

dat = dat %>% mutate(y.mod=mod$summary.fitted.values$mean[fitted.idx], 
                     i0.mu = mod$summary.random$i0$mean[fitted.idx], 
                     i1.mu = mod$summary.random$i1$mean[fitted.idx])
```

## Adaptations

I'm adapting the example to better align it with my particular data to see if I can still model and understand it.

Generate random spatial locations
```{r}
n <- 150
set.seed(1)
coo <- matrix(runif(2 * n), n)
```

Generate spatially and temporally correlated data.
The differences here with the original are (1) we are focused on the coefficient with respect to time (rather than `hh`) and (2) the time coefficient only varies spatially (with no need therefore of a temporal autocorrelation component for it). 
```{r}
kappa <- c(10, 12)
sigma2 <- c(0.75, 1.1)
k <- 15
rho <- c(0.7, 0.5)

set.seed(2)
beta0 <- book.rMatern(k, coo, range = sqrt(8) / kappa[1],
  sigma = sigma2[1])

set.seed(3)
beta1 <- book.rMatern(1, coo, range = sqrt(8) / kappa[2],
  sigma = sigma2[2])
beta0[, 1] <- beta0[, 1] / (1 - rho[1]^2)
# beta1[, 1] <- beta1[, 1] / (1 - rho[2]^2)

for (j in 2:k) {
  beta0[, j] <- beta0[, j - 1] * rho[1] + beta0[, j] *
    (1 - rho[1]^2)
  # beta1[, j] <- beta1[, j - 1] * rho[2] + beta1[, j] *
  #   (1 - rho[2]^2)
}

# Remove temporal dependence of beta1 through averaging (each year has same value for beta1)
# beta1 = as.vector(rowMeans(beta1))

set.seed(4)
# Simulate the covariate values
hh <- runif(n * k) 
mu.beta <- c(-5, 1)
taue <- 20

set.seed(5)
# Error in the observation
error <- rnorm(n * k, 0, sqrt(1 / taue))

# DGK Model year as covariate
Year = matrix(data=(1:k), nrow=n, ncol=k, byrow=TRUE)
y = (mu.beta[1] + beta0) + (mu.beta[2] + beta1) * Year +
  error
# year.val = matrix(data=(1:k)/k, nrow=n, ncol=k, byrow=TRUE)
# y = (mu.beta[1] + beta0) + (mu.beta[2] + beta1) * year.val +
#   error

# Arrange in tbl
dat = tibble(loc.idx=rep(1:n,k), E=rep(coo[,1],k), N=rep(coo[,2],k), 
             Year=rep(1:k, each=n), y=c(y), yr.val = c(year.val))
```

Exploration of the data.
```{r}
yr.trends = dat %>% 
  group_by(loc.idx) %>% 
  do(tidy(lm(y ~ Year, data = .))) %>% 
  filter(term=='Year') %>% select(-term)
dat.mu = dat %>% group_by(loc.idx) %>% 
  summarize(E=mean(E), N=mean(N), y=mean(y)) %>% 
  left_join(yr.trends %>% select(loc.idx, estimate))

ggplot(dat.mu) + geom_point(aes(x=E, y=N, color=y)) + scale_color_viridis_c()
ggplot(dat.mu) + geom_point(aes(x=E, y=N, color=estimate))
ggplot(dat %>% group_by(loc.idx)) + geom_line(aes(x=Year, y=y, group=loc.idx), alpha=0.1)
```

```{r}
mesh <- inla.mesh.2d(dat %>% select(E,N), max.edge = c(0.15, 0.3),
  offset = c(0.05, 0.3), cutoff = 0.07)

plot(mesh)
points(dat %>% select(E, N), col = "red")

spde <- inla.spde2.pcmatern(mesh = mesh,
  prior.range = c(0.05, 0.01), # P(practic.range < 0.05) = 0.01
  prior.sigma = c(1, 0.01)) # P(sigma > 1) = 0.01

i0 <- inla.spde.make.index('i0', spde$n.spde, n.group = length(unique(dat$Year)))
i1 <- inla.spde.make.index('i1', spde$n.spde)

A0 <- inla.spde.make.A(mesh, loc=cbind(dat$E, dat$N), group = dat$Year)
A1 <- inla.spde.make.A(mesh, loc=cbind(dat$E, dat$N), weights = dat$Year)

dat.stack <- inla.stack(
  data = list(y = dat$y), 
  A = list(A0, A1, 1), 
  effects = list(i0, 
                 i1, 
                 tibble(Intercept = 1, mu1 = 1, 
                        Year=dat$Year)),
  tag = 'dat')

form <- y ~ 0 + mu1 + Year + # to fit mu_beta
  f(i0, model = spde, group = i0.group, 
    control.group = list(model = 'ar1')) + 
  f(i1, model = spde)
```

```{r}
mod.2 = inla(form, family = 'gaussian',
           data = inla.stack.data(dat.stack), 
           control.predictor = list(A = inla.stack.A(dat.stack)), 
           control.inla = list(int.strategy = 'eb'), # no integr. wrt theta)
           control.compute = list(config=TRUE))
```

```{r}
# i=sample.int(max(dat$loc.idx), size=1)
# ggplot(dat %>% filter(loc.idx==i)) + geom_line(aes(x=Year, y=y))
```

Posterior checks on the linear coefficient (combined global and spatial components) for time.
```{r}
# Posterior draws for year coefficients
draws=500
n.locs = spde$n.spde
tmp = inla.posterior.sample(draws, result = mod.2)
tmp2 = inla.posterior.sample.eval(c("Year", "i1"), tmp)
tmp.glob = unname(tmp2[1,])
tmp.i = unname(tmp2[2:(n.locs+1),])
tmp.tot = tmp.glob + tmp.i

coeff.ppc = tibble(Sample=rep(1:draws, each=n.locs), coeff=as.vector(tmp.tot))

ggplot(coeff.ppc %>% group_by(Sample)) + 
  geom_line(aes(x=coeff, group=Sample), stat="density", alpha=0.1) + 
  geom_density(data=dat.mu, aes(x=estimate), color='red') 
```