### 2022-06-01 notebooks/kronecker-tests.Rmd
Adds notebook for experimenting with kronecker products for st modeling

---
title: "Testing spacetime model with Kronecker sums on WAIS data"
output: html_notebook
---

This attempts to model the time coefficients of accumulation data following the example found in the [section on dynamic space-time regression](https://becarioprecario.bitbucket.io/spde-gitbook/ch-stapp.html#dynamic-space-time-regression) in "Advanced Spatial Modeling with Stochastic Partial Differential Equations Using R and INLA".
I adapt this to my particular use case and will eventually further modify it to use a Gamma distribution with log link.

```{r}
library(here)
library(INLA)
library(dplyr)
library(tidyr)
library(broom)
library(ggplot2)
library(spdep)
```

```{r}
data = readRDS(here('data/Rdata-clean.rds'))
gdf_traces = readRDS(here('data/Rdata-gdf_trace.rds'))

set.seed(777)
gdf_traces = gdf_traces %>% sample_frac(0.01)

# # Subset to smaller region
# tmp = st_bbox(gdf_traces)
# East = c(tmp[1] + (tmp[3]-tmp[1])/7, tmp[1] + (tmp[3]-tmp[1])/2)
# North = c(tmp[2] + (tmp[4]-tmp[2])/2, tmp[4] - (tmp[4]-tmp[2])/8)
# tmp.df = tibble(East=East, North=North)
# poly = tmp.df %>%
#   st_as_sf(coords = c("East", "North"),
#            crs = 3031) %>%
#   st_bbox() %>%
#   st_as_sfc()
# gdf.idx = sapply(st_intersects(gdf_traces, poly),function(x){length(x)==0})
# gdf_traces = gdf_traces[!gdf.idx,]

data = data %>% filter(trace_ID %in% gdf_traces$trace_ID) %>% 
  filter(Year >= 1979) %>% filter(Year < 2010) %>% arrange(trace_ID, Year)

# Select variables of interest
dat = data %>% select(trace_ID, East, North, Year, accum, 
                      std, elev, dem, slope, aspect, 
                      # u10m, v10m, mu.u10, mu.v10
                      )

# Center covariates
dat = dat %>% 
  mutate(elev = elev-mean(elev, na.rm=TRUE), 
         dem=dem-mean(dem, na.rm=TRUE), slope=slope-mean(slope, na.rm=TRUE), 
         # u10m=u10m-mean(u10m, na.rm=TRUE), v10m=v10m-mean(v10m, na.rm=TRUE), 
         # mu.u10=mu.u10-mean(mu.u10, na.rm=TRUE), mu.v10=mu.v10-mean(mu.v10, na.rm=TRUE)
         ) %>% 
  mutate(Year.mod = Year-mean(Year), Year.idx = (Year-min(Year)+1))
         # u10m=mean(u10m, na.rm=TRUE), v10m=mean(v10m, na.rm=TRUE))

# Scale covariates by sd
dat = dat %>% 
  mutate(elev = elev/sd(elev), dem=dem/sd(dem), 
         slope=slope/sd(slope), 
         # u10m=u10m/sd(u10m), v10m=v10m/sd(v10m), mu.u10=sd(mu.u10), mu.v10=sd(mu.v10)
         )

# Split into training and testing sets
dat = dat %>% mutate(row.ID = 1:nrow(dat)) %>% relocate(row.ID)
dat.train = dat %>% slice_sample(prop = 0.80) %>% arrange(row.ID)
dat.test = dat %>% filter(!row.ID %in% dat.train$row.ID)
```

```{r}
# More coarse mesh for development purposes
mesh = inla.mesh.2d(loc = dat.train %>% select(East, North), 
                    max.edge = c(60000, 120000), cutoff = 12000)
# mesh = inla.mesh.2d(loc = dat.train %>% select(East, North), 
#                     max.edge = c(30000, 90000), cutoff = 1000)
plot(mesh)
points(dat.train %>% select(East, North), col = "red")

# These values will also need to be adjusted when going back to Gamma?
spde = inla.spde2.pcmatern(mesh = mesh, 
  prior.range = c(20000, 0.01), # P(range < 15 km) = 0.01
  prior.sigma = c(100, 0.05))

st.idx <- inla.spde.make.index('st.idx', spde$n.spde, n.group = length(unique(dat.train$Year.idx)))
yr.idx <- inla.spde.make.index('yr.idx', spde$n.spde)

A.st <- inla.spde.make.A(mesh, 
                         loc=cbind(dat.train$East, dat.train$North), 
                         group = dat.train$Year.idx)
A.yr <- inla.spde.make.A(mesh, 
                         loc=cbind(dat.train$East, dat.train$North), 
                         weights = dat.train$Year.mod)

dat.stack <- inla.stack(
  data = list(y = dat.train$accum), 
  A = list(1, 1, A.st, A.yr),
  effects = list(list(Intercept = rep(1, nrow(dat.train))), 
                 tibble(elev = dat.train$elev, dem=dat.train$dem, 
                        # slope=dat.train$slope, aspect=dat.train$aspect, 
                        Year=dat.train$Year.mod),
                 st.idx, 
                 yr.idx),
  tag = 'dat')

# Model formula
form = y ~ -1 + Intercept + #modeling the intercept
  # dem + aspect + #fixed effects
  f(st.idx, model = spde, group=st.idx.group, 
    control.group=list(model='ar1')) + #spatial random effect with AR1 temporal autocorrelation
  f(yr.idx, model = spde) #Global effect from year and deviation from global (modeled as a spatial random effect with weights given by Year)
```

## Gamma modeling

```{r mod-gamma, cache=TRUE}
# For testing, initially use gaussian

# mod.gamma = inla(form,
#                data = inla.stack.data(dat.stack),
#                family = 'Gamma',
#                control.predictor = list(compute = TRUE, A = inla.stack.A(dat.stack), link=1),
#                control.inla = list(int.strategy='eb', strategy = 'gaussian'), # Options improve computation time at cost of accuracy/precision
#                control.compute = list(waic=TRUE, config=TRUE))
mod.gamma = readRDS(here('data/interim_results/mod-gamma.rds'))

summary(mod.gamma)
```

### Comparisons to GLM models

```{r}
f.glm = y ~ -1 + Intercept + Year

glm.gauss = glm(f.glm, data = dat.train %>% rename(y=accum) %>% mutate(Intercept=1, Year=Year.mod))
glm.logGaus = mod.glm = glm(f.glm, data = dat.train %>% rename(y=accum) %>% mutate(Intercept=1, Year=Year.mod), 
               family = gaussian(link = "log"))
glm.gamma = glm(f.glm, data = dat.train %>% rename(y=accum) %>% mutate(Intercept=1, Year=Year.mod), 
               family = Gamma(link = "log"))
summary(glm.gamma)
```

### Posterior checks

Directly calculate linear coefficients of time from raw data.
```{r}
yr.trends = dat.train %>% 
  group_by(trace_ID) %>% 
  do(tidy(lm(accum ~ Year.mod, data = .))) %>% 
  filter(term=='Year.mod') %>% select(-term)
dat.mu = dat.train %>% group_by(trace_ID) %>% 
  summarize(East=mean(East), North=mean(North), accum=mean(accum)) %>% 
  left_join(yr.trends %>% select(trace_ID, estimate)) %>% 
  mutate(log.est = log(1+(estimate/accum)))
```

```{r}
# Posterior draws for year coefficients
draws=100
n.locs = spde$n.spde
mod.samples = inla.posterior.sample(draws, result = mod.gamma)
tmp = inla.posterior.sample.eval(c("yr.idx"), mod.samples)
coeff.ppc = tibble(Sample=rep(1:draws, each=n.locs), coeff=as.vector(tmp))

ggplot(coeff.ppc %>% group_by(Sample)) + 
  geom_line(aes(x=coeff, group=Sample), stat="density", alpha=0.1) + 
  geom_density(data=dat.mu, aes(x=log.est), color='red') + 
  xlim(c(-0.03, .03))
```

### Spatial comparisons

```{r}
tmp = tibble(Source="model", 
             East=mesh$loc[,1], North=mesh$loc[,2], 
             med=mod.gamma$summary.random$yr.idx$`0.5quant`, 
             LB=mod.gamma$summary.random$yr.idx$`0.025quant`, 
             UB=mod.gamma$summary.random$yr.idx$`0.975quant`) %>% 
  # mutate(weak=(LB<0 && UB>0)) %>% 
  st_as_sf(coords=c("East", "North"), crs=3031)

trends.sf = dat.mu %>% select(-trace_ID, -accum, -estimate)  %>% 
  rename(med=log.est) %>% mutate(LB=NA, UB=NA, Source="data") %>% 
  st_as_sf(coords=c("East", "North"), crs=3031) %>%
  bind_rows(tmp)

ggplot(trends.sf %>% filter(Source=="data")) + 
  geom_sf(aes(color=med)) + scale_color_gradient2(limits=c(-0.03, 0.03))
ggplot(trends.sf %>% filter(Source=="model")) + 
  geom_sf(aes(color=LB)) + scale_color_gradient2(limits=c(-0.03, 0.03))
ggplot(trends.sf %>% filter(Source=="model")) + 
  geom_sf(aes(color=med)) + scale_color_gradient2(limits=c(-0.03, 0.03))
ggplot(trends.sf %>% filter(Source=="model")) + 
  geom_sf(aes(color=UB)) + scale_color_gradient2(limits=c(-0.03, 0.03))
```
## Spacetime random effect removed

It's possible that I am removing temporal signal from my estimates when I model the spacetime auto-correlations.
Here I run the model without this random effect for comparison (so only using a spatially varying temporal coefficient).

```{r}
# Model formula
f.noST = y ~ -1 + Intercept + 
  # f(st.idx, model = spde, group=st.idx.group, 
  #   control.group=list(model='ar1')) + 
  f(yr.idx, model = spde)

mod.noST = inla(f.noST,
                data = inla.stack.data(dat.stack),
                family = 'Gamma',
                control.predictor = list(compute = TRUE, A = inla.stack.A(dat.stack), link=1),
                control.inla = list(int.strategy='eb', strategy = 'gaussian'), # Options improve computation time at cost of accuracy/precision
                control.compute = list(waic=TRUE, config=TRUE))

summary(mod.noST)
```

### Posterior checks

```{r}
# Posterior draws for year coefficients
draws=100
n.locs = spde$n.spde
samples.noST = inla.posterior.sample(draws, result = mod.noST)
tmp = inla.posterior.sample.eval(c("yr.idx"), samples.noST)
ppc.noST = tibble(Sample=rep(1:draws, each=n.locs), coeff=as.vector(tmp))

ggplot(ppc.noST %>% group_by(Sample)) + 
  geom_line(aes(x=coeff, group=Sample), stat="density", alpha=0.1) + 
  geom_density(data=dat.mu, aes(x=log.est), color='red') + 
  xlim(c(-0.03, .03))
```

### Spatial comparisons

```{r}
tmp = tibble(Source="model", 
             East=mesh$loc[,1], North=mesh$loc[,2], 
             med=mod.noST$summary.random$yr.idx$`0.5quant`, 
             LB=mod.noST$summary.random$yr.idx$`0.025quant`, 
             UB=mod.noST$summary.random$yr.idx$`0.975quant`) %>% 
  # mutate(weak=(LB<0 && UB>0)) %>% 
  st_as_sf(coords=c("East", "North"), crs=3031)

trends.noST = dat.mu %>% select(-trace_ID, -accum, -estimate)  %>% 
  rename(med=log.est) %>% mutate(LB=NA, UB=NA, Source="data") %>% 
  st_as_sf(coords=c("East", "North"), crs=3031) %>%
  bind_rows(tmp)

ggplot(trends.noST %>% filter(Source=="data")) + 
  geom_sf(aes(color=med)) + scale_color_gradient2(limits=c(-0.03, 0.03))
ggplot(trends.noST %>% filter(Source=="model")) + 
  geom_sf(aes(color=LB)) + scale_color_gradient2(limits=c(-0.03, 0.03))
ggplot(trends.noST %>% filter(Source=="model")) + 
  geom_sf(aes(color=med)) + scale_color_gradient2(limits=c(-0.03, 0.03))
ggplot(trends.noST %>% filter(Source=="model")) + 
  geom_sf(aes(color=UB)) + scale_color_gradient2(limits=c(-0.03, 0.03))
```

## Lognormal modeling

```{r mod-gauss, cache=TRUE}

# stack.LN <- inla.stack(
#   data = list(y = log(dat.train$accum)), 
#   A = list(1, 1, A.st, A.yr),
#   effects = list(list(Intercept = rep(1, nrow(dat.train))), 
#                  tibble(elev = dat.train$elev, dem=dat.train$dem, 
#                         # slope=dat.train$slope, aspect=dat.train$aspect, 
#                         Year=dat.train$Year.mod),
#                  st.idx, 
#                  yr.idx),
#   tag = 'dat')
# 
# # For testing, initially use gaussian
# mod.gauss = inla(form, 
#                data = inla.stack.data(stack.LN), 
#                family = 'Gaussian',
#                control.predictor = list(compute = TRUE, A = inla.stack.A(stack.LN)), 
#                control.compute = list(waic=TRUE, config=TRUE))
# 
# # mod.gauss = inla(form, 
# #                  data = inla.stack.data(dat.stack), 
# #                  )
# 
# summary(mod.gauss)
```