### 2022-07-11 notebooks/exploratory.Rmd

---
title: "WAIS analysis using INLA"
output: html_notebook
---

This is a working version of the general workflow notebook for processing WAIS radar results using INLA.
This includes the general process from raw data import to final processing.
This also includes some instructions for various pre-processing in Python.

## MISC USEFUL LINKS

A few other useful tutorials and such regarding R-INLA and INLA in general:

- [Official R-INLA Website](https://www.r-inla.org/)
- [Spatial Modeling with R-INLA](https://ourcodingclub.github.io/tutorials/spatial-modelling-inla/): A nice walkthrough tutorial of spatial modeling and geostatistics using r-inla
- [Book: Spatiotemporal Bayesian Models using INLA](https://sites.google.com/a/r-inla.org/stbook/home)
- [A gentle INLA tutorial](https://www.precision-analytics.ca/articles/a-gentle-inla-tutorial/)
- [Intro in INLA for geospatial modeling](https://punama.github.io/BDI_INLA/)
- [Digital Book: Bayesian inference with INLA](https://becarioprecario.bitbucket.io/inla-gitbook/index.html)
- [Brief intro/tutorial on INLA](https://www.flutterbys.com.au/stats/tut/tut12.9.html)
- [Great mini tutorials using R-INLA](https://haakonbakkagit.github.io/organisedtopics.html)
- [Book: Geospatial Health Data: Modeling and Visualization with R-INLA and Shiny](https://www.paulamoraga.com/book-geospatial/): Walkthroughs using health data, but the explanations make more intuitive sense to me. Section 8 is particularly useful for my work (sadly Section 10 appears to be blank)

## Python preprocessing

To get data in the proper format prior to these analyses, some Python processing is required.
The necessary data are produced and saved to disk by running `preprocessing.py` in the `src` directory.

## INLA modeling

The remainder of this notebook shows the workflow for processing PAIPR WAIS data using rINLA.
I first detail some of the theory and principles behind this, and how specifically I plan to utilize them in this research specifically.

### Background

The accumulation data modeled in this study represents a continuous distribution of a random variable over two continuous spatial dimensions and indexed over an additional discrete temporal dimension.
We model accumulation as a gamma distribution to ensure a zero-bounded function with increased variance at higher expected values.
$$\mathbf{Y} \sim Gamma(a,b)$$
Rather than using the shape and rate parameters $a$ and $b$, we can express this gamma distribution in terms of the expected y value $E(y_i) = a_i/b_i = \mu_i$ and its associated variance (as expressed using the precision parameter $\phi$) $Var(y_i) = a_i/b_i^2 = \mu_i^2/\phi$.
This $\phi$ is the same precision parameter as output from `inla` (there is also a scale parameter $s$ that has a default of 1 where $Var(y_i) = 1/\tau$ and $\tau = (s\phi)/\mu_i^2$).
With rearranging and substituting, we have $a_i = \phi$ and $b_i = \phi/\mu_i$ and we can therefore express the distribution in accumulation observations in terms of $\phi$ and $\mu_i$.
$$y_i \sim Gamma(\mu_i, \phi)$$

For each observation $y_i \in \mathbf{Y}$, we can construct a linear predictor $\eta_i$ for the expectation value $\mu_i$.
As we assume a gamma-distribution for accumulation, we associate $\eta_i$ and $\mu_i$ via the log link function.
$$\log(\mu_i) = \eta_i$$ 
We can further model $\eta_i$ as a generalized linear model consisting of fixed effect variables $\mathbf{X}$ and smoothed random latent effects (in this case we are modeling a spatial random effect $u_i$ and a temporal random effect $\omega_i$).
$$
\eta_{i} = \alpha +  \sum^M_{m=1} \beta_m x_{mi} + u_i + \omega_{i}
$$

These latent random effects $u_i$ and $\omega_i$ are indexed by a set of parameters $\theta_s$ and $\theta_t$ that account for the spatial and temporal correlation in the data.
We can model these correlation structures (as defined by the parameters $\pmb{\theta}$) as latent stationary Gaussian Fields, using a function of some hyperparameters $\psi$ and an associated prior distribution $p(\psi)$.
This is equivalent to assuming that $\pmb{\theta}$ is associated with some multivariate Normal distribution with mean $\mu = (\mu_1,...,\mu_n)'$ and covariance structure $\Sigma$, where $\Sigma_{ij} = Cov(\theta_i, \theta_j)$.

$$
\theta \sim \mathcal{N}(\mu, \Sigma)
$$

If we assume adherence to Markovian properties, our precision matrix $Q = \Sigma^{-1}$ is sparse, with the non-zero components of $Q$ completely given by the neighborhood structure ($N$) of our process, i.e. $Q_{ij} \neq 0 \iff j \in \{i,N(i)\}$.
In this assumption, we specify a Gaussian-Markov Random Field for both our spatial and temporal structures, permitting vast improvements in computational efficiency.
We also model the spatial and temporal random effects independently, which again greatly improves computation time, but will fail to capture any co-dependencies between these effects.

The spatial GMRF $u_i$ is derived from a stochastic partial differential equation...

We model the temporal random effect $\omega_i$ as a first-order autoregressive model^[I will eventually want to update this to an AR(2) process due to the known temporal autocorrelation in our data] (representing how the spatial field evolves in time) where
$$\omega_t = \rho \omega_{t-1}$$

Our objective, therefore, is to find the joint posterior distribution
$$P(\phi, \alpha, \pmb{\beta}, \rho, \mid \mathbf{Y})$$

### INLA installation

The latest version of R (`4.1.1` as of this writing) states it is currently incompatible with R-INLA.
This seems to be some sort of bug (perhaps limited to RStudio?) as it is possible to do this.
It does require installing from a local source instead of following the [standard install instructions](https://www.r-inla.org/download-install) for R-INLA.
Binaries for R-INLA can be found at [this ftp](http://inla.r-inla-download.org/R/testing/src/contrib/).
Then simply follow [these instructions](https://riptutorial.com/r/example/5556/install-package-from-local-source) to do install from local source in RStudio.
This will also require installing some of the dependencies and misc. required R packages as well (e.g. `sp`, `foreach`, etc.).

### Imports

```{r}
library(here)
library(tidyverse)
library(cetcolor)
library(sf)
library(terra)
library(INLA)
library(INLAutils)
library(spdep)
```

### Data loading and pre-processing

```{r}
data_long = read_csv(here('data/paipr-long.csv')) %>% select(-...1)
gdf_traces = st_read(here('data/traces.geojson')) %>% as_tibble() %>% st_as_sf()

# Remove points in far northern extent
accum_bnds = st_coordinates(gdf_traces)
gdf_traces = gdf_traces %>% st_crop(xmin=-1.5e6, xmax=max(accum_bnds[,1]),
                                    ymin=min(accum_bnds[,2]), ymax=max(accum_bnds[,2]))

# Remove unrealistically high elevations
trace_drop = gdf_traces %>% filter(elev > 3000)

# Filter out overly large accum sites
trace_drop = trace_drop %>% filter(accum > 1000) %>% select(trace_ID)

# Add East/North coordinates and drop large accum sites
pts_tmp = st_coordinates(gdf_traces)
coord_df = tibble(trace_ID = gdf_traces$trace_ID, East = pts_tmp[,1], North = pts_tmp[,2])
data = data_long %>% left_join(coord_df) %>% filter(!trace_ID %in% trace_drop$trace_ID)
```

For testing/development purposes, only use a small sample of total
```{r}
# xy.bnds = st_bbox(gdf_traces)
# xy.range = list(x=(xy.bnds$xmax - xy.bnds$xmin), y=(xy.bnds$ymax - xy.bnds$ymin))
# gdf_traces = gdf_traces %>% st_crop(xmin=as.numeric(xy.bnds$xmin+0.25*xy.range$x), 
#                                     xmax=as.numeric(xy.bnds$xmin+0.50*xy.range$x),
#                                     ymin=as.numeric(xy.bnds$ymin), 
#                                     ymax=as.numeric(xy.bnds$ymin+0.25*xy.range$y))

set.seed(777)
gdf_traces = gdf_traces %>% sample_frac(0.01)
data = data %>% filter(trace_ID %in% gdf_traces$trace_ID) %>% 
  filter(Year >= 1980) %>% filter(Year <= 2000) %>% arrange(trace_ID, Year)
```

```{r}
# Select variables of interest
dat = data %>% select(trace_ID, East, North, Year, accum, std, elev)

# Center covariates
yr.min = min(dat$Year)-1
dat = dat %>% mutate(Year = Year-yr.min, elev = elev-mean(elev))

# Split into training and testing sets
dat = dat %>% mutate(row.ID = 1:nrow(dat)) %>% relocate(row.ID)
dat.train = dat %>% slice_sample(prop = 0.80) %>% arrange(row.ID)
dat.test = dat %>% filter(!row.ID %in% dat.train$row.ID)
```

### INLA  modeling

We first create the mesh used for defining the neighborhoods (this is important for generating the sparse precision matrix $Q$, enabling our GMRF approximation).
```{r}
mesh = inla.mesh.2d(loc = dat.train %>% select(East, North), 
                    max.edge = c(20000, 50000), cutoff = 1000)
plot(mesh)
points(dat.train %>% select(East, North), col = "red")
# mesh = inla.mesh.2d(loc = st_coordinates(gdf_traces), 
#                     max.edge = c(20000, 50000), cutoff = 1000)
# plot(mesh)
# points(st_coordinates(gdf_traces), col = "red")
```

<!-- ## Form 1 -->

<!-- This follows examples as shown in Section 7.1 of [Advanced Spatial Modeling with Stochastic Partial Differential Equations Using R and INLA](https://becarioprecario.bitbucket.io/spde-gitbook/ch-spacetime.html#discrete-time-domain). -->

<!-- ### -->

<!-- We then define our SPDE object and create the identifier matrix $A$. -->
<!-- ```{r} -->
<!-- spde = inla.spde2.pcmatern(mesh = mesh,  -->
<!--   prior.range = c(5000, 0.01), # P(range < 5 km) = 0.01 -->
<!--   prior.sigma = c(250, 0.1)) # P(sigma > 250 mm w.e.) = 0.10 -->
<!-- # spde = inla.spde2.matern(mesh=mesh, alpha=2) -->

<!-- iset = inla.spde.make.index('i', n.spde = spde$n.spde, -->
<!--   n.group = max(dat$Year)) -->
<!-- A = inla.spde.make.A(mesh = mesh, -->
<!--   loc = cbind(dat$East, dat$North), group = dat$Year)  -->
<!-- # A_matrix = inla.spde.make.A(mesh, loc=st_coordinates(gdf_traces)) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- dat.stack = inla.stack( -->
<!--   data = list(y = dat$accum),  -->
<!--   A = list(A, 1),  -->
<!--   effects = list(iset, data.frame(Intercept = 0, elev = dat$elev)), -->
<!--   tag = 'dat')  -->
<!-- ``` -->

<!-- I also define a PC-prior for the autoregressive temporal autocorrelation. -->
<!-- In particular, I assume that $P(corr > 0) = 0.9$, indicating a high degree of confidence that some temporal autocorrelation exists in our data. -->
<!-- ```{r} -->
<!-- # Prior on autoregressive rho parameter -->
<!-- h.spec <- list(rho = list(prior = 'pc.cor1', param = c(0, 0.9))) -->

<!-- # Prior on fixed effects intercept -->
<!-- # prior.fixed <- list(mean.intercept = , prec.intercept = ) -->

<!-- # Model formula -->
<!-- formula = y ~ 0 + Intercept + elev + f(i, model = spde, group = i.group,  -->
<!--   control.group = list(model = 'ar1', hyper = h.spec)) -->

<!-- # PC prior on the autoregressive param -->
<!-- prec.prior <- list(prior = 'pc.prec', param = c(50, 0.9)) -->

<!-- # Model fitting -->
<!-- system.time( -->
<!--   results <- inla(formula,  -->
<!--                family = 'Gamma', -->
<!--                data = inla.stack.data(dat.stack),  -->
<!--                control.predictor = list(compute = TRUE, A = inla.stack.A(dat.stack)),  -->
<!--                # control.fixed = prior.fixed,  -->
<!--                control.family = list(hyper = list(prec = prec.prior)),  -->
<!--                control.compute = list(dic = TRUE)) -->
<!-- ) -->
<!-- ``` -->

<!-- ## Form 2 -->

<!-- This appears to be the same end result as in Form 1, but a slightly different implementation as followed by Section 8.6.2 in [Bayesian inference with INLA](https://becarioprecario.bitbucket.io/inla-gitbook/ch-temporal.html#separable-models). -->

<!-- ### -->

<!-- We then define our SPDE object and create the identifier matrix $A$. -->
<!-- ```{r} -->
<!-- spde = inla.spde2.pcmatern(mesh = mesh,  -->
<!--   prior.range = c(5000, 0.01), # P(range < 5 km) = 0.01 -->
<!--   prior.sigma = c(250, 0.1)) # P(sigma > 250 mm w.e.) = 0.10 -->
<!-- # spde = inla.spde2.matern(mesh=mesh, alpha=2) -->

<!-- iset = inla.spde.make.index('i', n.spde = spde$n.spde, -->
<!--   n.group = max(dat$Year)) -->
<!-- A = inla.spde.make.A(mesh = mesh, -->
<!--   loc = cbind(dat$East, dat$North), group = dat$Year)  -->
<!-- # A_matrix = inla.spde.make.A(mesh, loc=st_coordinates(gdf_traces)) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- dat.stack = inla.stack( -->
<!--   data = list(y = dat$accum),  -->
<!--   A = list(A, 1),  -->
<!--   effects = list(iset, data.frame(Intercept = 0, elev = dat$elev)), -->
<!--   tag = 'dat')  -->
<!-- ``` -->

<!-- I also define a PC-prior for the autoregressive temporal autocorrelation. -->
<!-- In particular, I assume that $P(corr > 0) = 0.9$, indicating a high degree of confidence that some temporal autocorrelation exists in our data. -->
<!-- ```{r} -->
<!-- # Prior on autoregressive rho parameter -->
<!-- h.spec <- list(rho = list(prior = 'pc.cor1', param = c(0, 0.9))) -->

<!-- # Prior on fixed effects intercept -->
<!-- # prior.fixed <- list(mean.intercept = , prec.intercept = ) -->

<!-- # Model formula -->
<!-- formula = y ~ 0 + Intercept + elev + f(i, model = spde, group = i.group,  -->
<!--   control.group = list(model = 'ar1', hyper = h.spec)) -->

<!-- # PC prior on the autoregressive param -->
<!-- prec.prior <- list(prior = 'pc.prec', param = c(50, 0.9)) -->

<!-- # Model fitting -->
<!-- system.time( -->
<!--   results <- inla(formula,  -->
<!--                family = 'Gamma', -->
<!--                data = inla.stack.data(dat.stack),  -->
<!--                control.predictor = list(compute = TRUE, A = inla.stack.A(dat.stack)),  -->
<!--                # control.fixed = prior.fixed,  -->
<!--                control.family = list(hyper = list(prec = prec.prior)),  -->
<!--                control.compute = list(dic = TRUE)) -->
<!-- ) -->
<!-- ``` -->

## Form 3

This follows examples from Section 8.6.1 in [Bayesian inference with INLA](https://becarioprecario.bitbucket.io/inla-gitbook/ch-temporal.html#separable-models), which details the spatio-temporal component as a separable model more explicitly (two random effects in the formula, one for the spatial model and one for the group-indexed time model).

**THIS IS SO MUCH FASTER!!!**

###

We then define our SPDE object and create the identifier matrix $A$.
```{r}
spde = inla.spde2.pcmatern(mesh = mesh, 
  prior.range = c(10000, 0.01), # P(range < 10 km) = 0.01
  prior.sigma = c(250, 0.10)) # P(sigma > 250 mm w.e.) = 0.10
# spde = inla.spde2.matern(mesh=mesh, alpha=2)

# iset = inla.spde.make.index('i', n.spde = spde$n.spde,
#   n.group = max(dat$Year))

# Create projector matrix for SPDE mesh
A.spat = inla.spde.make.A(mesh = mesh, loc = cbind(dat.train$East, dat.train$North))
# A.spat = inla.spde.make.A(mesh = mesh, loc = cbind(dat$East, dat$North))

# Assign index vectors for SPDE model
spat.idx = inla.spde.make.index(name = "spat.field", n.spde = spde$n.spde)
```

```{r}
# Make data stack
dat.stack <- inla.stack(
  data = list(y = dat.train$accum), 
  A = list(1, 1, 1, A.spat),
  effects = list(list(Intercept = rep(1, nrow(dat.train))), 
                 tibble(elev = dat.train$elev), 
                 list(time = dat.train$Year), 
                 spat.idx),
  tag = 'dat') 
```

I also define a PC-prior for the autoregressive temporal autocorrelation.
In particular, I assume that $P(corr > 0) = 0.9$, indicating a high degree of confidence that some temporal autocorrelation exists in our data.
```{r}
## Define other priors (spatial field priors defined earlier)

# Prior on autoregressive rho parameter
time.spec <- list(rho = list(prior = 'pc.cor1', param = c(0, 0.9)))

# Prior on fixed effects intercept
prior.fixed <- list()

# Prior on precision parameter for Gamma observations
prec.prior <- list(prior = 'pc.prec', param = c(250, 0.1)) #P(sigma > 250) = 10%

## PC prior implementations

# # Prior on autoregressive rho parameter
# time.spec <- list(rho = list(prior = 'pc.cor1', param = c(0, 0.9)))
# 
# # Prior on fixed effects intercept
# prior.fixed <- list()
# 
# # Prior on precision parameter for Gamma observations
# prec.prior <- list(prior = 'pc.prec', param = c(250, 0.1)) #P(sigma > 250) = 10%

# Model formula
formula = y ~ -1 + Intercept + elev + 
  f(time, model = 'ar1', hyper = time.spec) + 
  f(spat.field, model = spde)
# formula = y ~ elev + f(spat.field, model = spde) + f(time.field, model = 'ar1', hyper = h.spec)
```

```{r}
# Model fitting
system.time(
  results <- inla(formula, 
                  data = inla.stack.data(dat.stack), 
                  family = 'Gamma',
                  control.predictor = list(compute = TRUE, A = inla.stack.A(dat.stack), link = 1), 
                  # control.fixed = prior.fixed, 
                  control.family = list(hyper = list(prec = prec.prior)), 
                  control.compute = list(dic = TRUE))
)
```

## Continuation

```{r}
# summary(results)
round(results$summary.fixed, 4)
round(results$summary.hyperpar, 3)
results$dic$dic
```

Extract fitted values from `inla` model.
```{r}
dat.idx = inla.stack.index(dat.stack, 'dat')$data
G.params = tibble(mu = results$summary.fitted.values$mean[dat.idx], mu.sd = results$summary.fitted.values$sd[dat.idx])

res.tbl = tibble(trace_ID = dat.train$trace_ID, East = dat.train$East, North = dat.train$North, 
              Year = dat.train$Year, elev = dat.train$elev, accum = dat.train$accum, y.hat = G.params$mu) %>% 
  mutate(y.res = y.hat-accum) %>% 
  st_as_sf(coords = c("East", "North"), crs=st_crs(gdf_traces))

```

## Test data results

```{r}
# Create projector matrix for test data
A.test = inla.spde.make.A(mesh = mesh, loc = cbind(dat.test$East, dat.test$North))

# Create test data stack and join to data stack
test.stack <- inla.stack(
  data = list(y = NA), 
  A = list(1, 1, 1, A.test),
  effects = list(list(Intercept = rep(1, nrow(dat.test))), 
                 tibble(elev = dat.test$elev), 
                 list(time = dat.test$Year), 
                 spat.idx),
  tag = 'test')
val.stack <- inla.stack(dat.stack, test.stack)
```

```{r}
# Model fitting to test data
system.time(
  mod.test <- inla(formula, 
                  data = inla.stack.data(val.stack), 
                  family = 'Gamma',
                  quantiles = NULL, 
                  control.predictor = list(compute = TRUE, A = inla.stack.A(val.stack), link = 1), 
                  control.family = list(hyper = list(prec = prec.prior)), 
                  control.inla = list(strategy = 'adaptive'), 
                  control.mode = list(theta = results$mode$theta, restart = FALSE))
)
```

Below is a function that, given an input of mean estimates, sd on mean estimates, precision parameter for Gamma observations, and sd on precision parameter, returns $n$ samples from the posterior predictive distribution for the given inputs.
```{r}
pp.dist = function(data, prec, prec.sd, n=5000, return_sample=FALSE) {
  if (return_sample==TRUE) {
    ppd = matrix(data=NA, nrow=nrow(data), ncol=n)
  } else {
    ppd = tibble(mu=rep(0,nrow(data)), map=rep(0,nrow(data)), 
                 CI.low=rep(0,nrow(data)), CI.high=rep(0,nrow(data)))
  }

  for (i in 1:nrow(data)) {
    dat.i = data[i,]
    mu = rnorm(n, mean=dat.i$mu, sd=dat.i$mu.sd)
    a = rnorm(n, mean=prec, sd=prec.sd)
    b = a/mu
    dist = rgamma(n=n, shape = a, rate = b)

    if (return_sample==TRUE) {
      ppd[i,] = dist
    } else {
      dist.mu = mean(dist)
      d = density(dist)
      dist.map = d$x[which.max(d$y)]
      dist.hdi = HDInterval::hdi(dist, credMass=0.89)
      ppd[i,] = tibble_row(mu=dist.mu, map=dist.map, 
                           CI.low=as.numeric(dist.hdi[1]), 
                           CI.high=as.numeric(dist.hdi[2]))
    }
  }
  return(ppd)
}
```

```{r}
# Get inla grid indices where test data resides
test.idx = inla.stack.index(val.stack, 'test')$data

test.fitted = tibble(mu = mod.test$summary.fitted.values$mean[test.idx], 
                                  mu.sd = mod.test$summary.fitted.values$sd[test.idx])

system.time(
  post.pred_test <- pp.dist(data=test.fitted, prec=mod.test$summary.hyperpar$mean[1], 
                       prec.sd=mod.test$summary.hyperpar$sd[1])
)
```

```{r}
res.test = dat.test %>% 
  mutate(mu=post.pred_test$mu, CI.low=post.pred_test$CI.low, 
                               CI.hi=post.pred_test$CI.high) %>% 
  mutate(y.res = mu-accum)

ggplot(data=res.test, aes(accum, mu)) + 
  geom_point() + 
  geom_abline(slope=1, intercept=0, color='red') + 
  # geom_smooth(method = 'lm') + 
  lims(x=c(200,600), y=c(200,600))

d = density(res.test$y.res)
d.mode = d$x[which.max(d$y)]
ggplot(res.test) + geom_density(aes(y.res)) + 
  geom_vline(xintercept = d.mode, linetype=2, color='red') + 
  annotate(geom="text", x=d.mode-100, y=max(d$y), 
           label=paste("MAP =", format(d.mode, digits=3)), 
           color="red")
```

###

Here's some experimental stuff for looking at the variability in our estimates.
Basically, I think this might represent the expected value (the `fitted.values` from the model) along with the model variance.

We have data drawn from a Gamma distribution
$$
y_i \mid \mathbf{X_i},\alpha,s_i,\theta \sim Gamma(a_i,b_i)
$$
where the expected value is $E(y_i) = a_i/b_i = \mu_i$ and the variance is $Var(y_i) = a_i/b_i^2 = \mu_i^2/\phi$ where $\phi$ is the precision parameter (this is the same precision parameter as output from `inla` - there is also a scale parameter that has a default of 1 where $Var(y_i) = 1/\tau$ and $\tau = (s\phi)/\mu_i^2$).
This in turn should mean that ??? $a_i = \phi$ and $b_i = \phi/\mu_i$ ???.

We can therefore report the calculated expected values $\mu_i \in \pmb{\mu}$ along with the modeled random variance $\pmb{\mu}^2/\phi$.
I'm not sure how this modeled random variance is supposed to differ from the sd reported in `fitted.values` though (I suppose sd must be the standard deviation on the mean parameter, while the modeled variance is the assumed variability about the true mean).