### 2022-07-20 notebooks/deprecated/article_draft.Rmd
Cleans up old notebook files

---
title: "Ch. 2 Dissertation draft"
csl: ../docs/american-geophysical-union.csl
link-citations: yes
output:
  html_document:
    df_print: paged
bibliography: ../docs/citations.bib
---

## README

This is a draft version of what will eventually become the paper I submit as the second chapter of my dissertation. 
Eventually it would be good for this to be a fully reproducible paper that I can submit to a journal, but that might require more $\LaTeX$ knowledge than I currently possess.

## Introduction

This is where the introduction will go.

## Methods

This is where the methods will go.

```{r, echo=FALSE, message=FALSE}

# Attach required packages
library(tidyverse)
library(here)
library(sf)
library(tcltk)
library(readxl)
library(tictoc)

tic("Total runtime")

# Import accumulation data results
accum.PS = read_rds(here("data/interim_results/accum_data.rds"))
cores.PS = read_rds(here("data/interim_results/core_data.rds"))

# # Code to export data as .csv for use in Python
# accum.exp = accum.PS %>% 
#   mutate(trace_ID = 0:(nrow(accum.PS)-1), Easting=st_coordinates(.)[,1], 
#          Northing=st_coordinates(.)[,2]) %>%
#   st_drop_geometry() %>% unnest(cols = data) %>% 
#   select(trace_ID, Easting, Northing, everything())
# cores.exp = cores.PS %>%
#   mutate(Easting=st_coordinates(.)[,1], Northing=st_coordinates(.)[,2]) %>%
#   st_drop_geometry() %>% unnest(cols = Data) %>% select(Site, Easting, Northing, everything())
# write_csv(accum.exp, '/home/durbank/Documents/Research/Antarctica/WAIS-wind/data/R-trends/accum.csv')
# write_csv(cores.exp, '/home/durbank/Documents/Research/Antarctica/WAIS-wind/data/R-trends/cores.csv')

# Filter accumulation data
accum.PS = accum.PS %>% filter(elev.REMA > 0) %>% 
  filter(slope < 2)

# Import basin shapefiles (in Antarctic Polar Sterographic EPSG:3031)
basins = st_read(here("data/ANT_Basins_IMBIE2_v1.6/ANT_Basins_IMBIE2_v1.6.shp"))

# Import Antarctic outline
Ant.outline = st_read(here("data/Ant_basemap/Coastline_medium_res_polygon.shp"))
```

Below is a reference map showing the location of our data within the greater context of Antarctica.

```{r, fig_inset, echo=FALSE}

# Plot of current Antarctic cores (blue) and soon-to-be-added (red)
cores.plt = ggplot(cores.PS) + geom_sf(data = Ant.outline, color = 'grey') +
  geom_sf(data = sample_n(accum.PS, size = 1000), color = 'red') + 
  geom_sf(aes(text=Site), color = 'blue') + coord_sf(datum = st_crs(cores.PS)) + 
  theme_void()
print(cores.plt)
```

## Results

This is where the results will go.

```{r, echo=FALSE, message=FALSE}

# Create non-referenced tibble of accumulation by location
accum.loc_PS = as_tibble(st_coordinates(accum.PS)) %>%
  rename(Easting = X, Northing = Y) %>% bind_cols(st_set_geometry(accum.PS, NULL))

# Create non-referenced tibble of core accumulation by location
cores.loc_PS = as_tibble(st_coordinates(cores.PS)) %>%
  rename(Easting = X, Northing = Y) %>% bind_cols(st_set_geometry(cores.PS, NULL))
```

```{r, accum_map, echo=FALSE}
# Plot mean accum on spatial grid
accum.plt1 = ggplot(accum.loc_PS, aes(x = Easting, y = Northing, color = accum_mu)) + 
  geom_point() + 
  geom_point(data = cores.loc_PS, aes(x = Easting, y = Northing, color = accum_mu), 
             shape = 12) + 
  scale_color_viridis_c() + 
  # ggtitle("Mean accum in space") + 
  xlab("Easting") + ylab("Northing") + theme_bw()
print(accum.plt1)
```

Above we see how mean accumulation rates vary over the region of interest.

```{r, trend_map, echo=FALSE}
# Plot scaled trends significant at 95% CL
trend.map = ggplot(accum.loc_PS, aes(x = Easting, y = Northing)) + 
  geom_point(aes(color = 100*coeff_perc)) + 
  geom_point(data = filter(accum.loc_PS, p.val_yr < 0.05) %>% sample_n(2000), 
             aes(x=Easting, y=Northing), shape = 4, color = 'black', size = 0.05) + 
  geom_point(data = cores.loc_PS, 
             aes(x = Easting, y = Northing, color = 100*coeff_perc), shape = 8, size = 3) + 
  geom_point(data = filter(cores.loc_PS, p.val_yr < 0.05), 
             aes(x=Easting, y=Northing), shape = 4, color = 'black', size = 0.05) + 
  lims(x = c(-1.5e6,-9e5), y = c(-6e5,0.35e5)) +
  scale_colour_gradient2() +
  labs(x = "Easting", y = "Northing", color = "%/yr") + 
  theme_bw()
print(trend.map)
```

Above we see how the linear time series trend in annual accumulation rate varies spatially across the region of interest.

```{r}
# test = filter(accum.loc_PS, Northing < -4.5E5 & Easting < -1.2E6)
# 
# test.smp = test %>% sample_n(size = 250)
# test.lm = map(test.smp$data, ~ lm(accum ~ Year, data = .x, weights = 1/sd^2))
# 
# test.data = dplyr::select(test.smp, -data, -accum_mu) %>% 
#   mutate(data = map(test.lm, 
#                     ~ dplyr::select(augment(.), Year, accum, .fitted, .se.fit, .resid, .std.resid)))
# 
# tmp.tidy = map(test.lm, ~ tidy(.x))
# test.smp = test.smp %>% 
#   mutate(coeff_yr = simplify(map(tmp.tidy, "estimate") %>% map(2))) %>% 
#   mutate(std.err = simplify(map(tmp.tidy, "std.error") %>% map(2))) %>% 
#   mutate(p.val_yr = simplify(map(tmp.tidy, "p.value") %>% map(2)))
# 
# 
# test2 = test.smp %>% sample_n(size = 10) %>% unnest(cols = c(data))
# 
# ggplot(test2, aes(Year, accum)) + geom_line(aes(group = Easting, color = factor(Easting)))

```

## Comparison of mean accumulation and slope

This section uses 8-meter mosaic tiles from the Reference Elevation Model of Antarctica (REMA) to compare mean accumulation rate to surface slope [@howat_reference_2019].

### Exploratory analyses

```{r, message=FALSE}
library(GGally)

ggpairs(select(accum.loc_PS, Easting, Northing, elev.REMA, slope, aspect, accum_mu))

accum_lm = lm(accum_mu ~ Easting + Northing + elev.REMA + slope + aspect, data = accum.loc_PS)
summary(accum_lm)

core.accum_lm = lm(accum_mu ~ Easting + Northing + Elevation, data = cores.loc_PS)

```

Although all of the included predictors show significant effects on `accum_mu`, the dominant influences on the region-wide trends are concentrated in the spatial components.
Interestingly, elevation shows residual influence on the predictions after accounting it's covariance in x-y space.
The following model summarizes these results for the dominant influences.

```{r}
summary(lm(accum_mu ~ Easting + Northing + elev.REMA, data = accum.loc_PS))
```

We can see that by excluding `slope` and `aspect` from the analysis the model remains virtually the same.
The coefficients for the remaining predictors are very similar as are the p-values.
The new $r^2 = 0.6453$ a neglibible difference from the first model.

The model suggests higher Easting and Northing values (in this case moving south and east respectively) leads to lower accumulation rates, which fits our intuition.
The model further suggests that, after accounting for the covariance of elevation with position, higher elevation sites experience higher accumulation rates. **[What does this mean?]**

### Spatial models

Just because Easting, Northing, and elevation dominate the regional trends does not mean slope and aspect are insignificant.
We therefore utilize spatial regression modeling with slope and aspect to investigate the more localized effects of these two predictors.

```{r, message=FALSE}

# test.data = st_crop(accum.PS, c(xmin=-1.4E6, xmax=(-1.4E6 + 50000), ymin=-5E5, ymax=-4.5E5))
test.data = filter(accum.loc_PS, Easting >= -1.3E6 & Easting <= (-1.3E6+30000)) %>% 
  filter(Northing >= -5E5 & Northing <= -4.5E5)
data.lm = lm(accum_mu ~ Easting + Northing + elev.REMA + slope + aspect, data=test.data)

summary(data.lm)

ggpairs(select(test.data, Easting, Northing, elev.REMA, slope, aspect, accum_mu))
ggplot(test.data, aes(x=Easting, y=Northing)) + 
  geom_point((aes(color=accum_mu))) + scale_color_viridis_c()
```

If we look at the residuals, most don't show additional signal, expect for the East/West spatial components. 
These still show cyclic variations in mean accumulation rates (I would assume lining up with the undulations expressed in the echograms).
The below plot shows residuals plotted against Easting as an example of this.

```{r}
plot(test.data$Easting, residuals(data.lm))
```

### Spatial analysis on individual years

In addition to looking at the mean aggregate accumulation relationships, I should also look at those relationships on the range of individual years (to determine the degree of noise injected by annual fluctuations).
Bear in mind that I will not have predictors at annual resolution.
This may not matter for some predictors (Easting, Northing, maybe elevation) but could be significant for others (slope and aspect?).

```{r}
toc()
```

# REFERENCES