### 2022-07-20 notebooks/deprecated/kronecker-tests.Rmd

---
title: "Testing spacetime model with Kronecker sums on WAIS data"
output: html_notebook
---

This attempts to model the time coefficients of accumulation data following the example found in the [section on dynamic space-time regression](https://becarioprecario.bitbucket.io/spde-gitbook/ch-stapp.html#dynamic-space-time-regression) in "Advanced Spatial Modeling with Stochastic Partial Differential Equations Using R and INLA".
I adapt this to my particular use case and will eventually further modify it to use a Gamma distribution with log link.

```{r}
library(here)
library(INLA)
library(dplyr)
library(tidyr)
library(broom)
library(ggplot2)
library(spdep)
```

```{r}
data = readRDS(here('data/Rdata-clean.rds'))
gdf_traces = readRDS(here('data/Rdata-gdf_trace.rds'))

# # Subset to smaller region
# tmp = st_bbox(gdf_traces)
# # East = c(tmp[1] + (tmp[3]-tmp[1])/7, tmp[1] + (tmp[3]-tmp[1])/2)
# # North = c(tmp[2] + (tmp[4]-tmp[2])/2, tmp[4] - (tmp[4]-tmp[2])/8)
# East = c(tmp[1] + (tmp[3]-tmp[1])/7, tmp[3])
# North = c(tmp[2], tmp[2] + (tmp[4]-tmp[2])/5)
# tmp.df = tibble(East=East, North=North)
# poly = tmp.df %>%
#   st_as_sf(coords = c("East", "North"),
#            crs = 3031) %>%
#   st_bbox() %>%
#   st_as_sfc()
# ggplot() + geom_sf(data=poly) + geom_sf(data=gdf_traces)
# gdf.idx = sapply(st_intersects(gdf_traces, poly),function(x){length(x)==0})
# gdf_traces = gdf_traces[!gdf.idx,]

set.seed(777)
# gdf_traces = gdf_traces %>% sample_frac(0.05)
skip.int = 5
gdf.idx = seq(1, nrow(gdf_traces), by=skip.int)
gdf_traces = gdf_traces[gdf.idx,]

data = data %>% filter(trace_ID %in% gdf_traces$trace_ID) %>% 
  filter(Year >= 1979) %>% filter(Year < 2010) %>% arrange(trace_ID, Year)

# Select variables of interest
dat = data %>% select(trace_ID, East, North, Year, accum, 
                      std, elev, dem, slope, aspect, 
                      # u10m, v10m, mu.u10, mu.v10
                      )

# Center covariates
dat = dat %>% 
  mutate(elev = elev-mean(elev, na.rm=TRUE), 
         dem=dem-mean(dem, na.rm=TRUE), slope=slope-mean(slope, na.rm=TRUE), 
         # u10m=u10m-mean(u10m, na.rm=TRUE), v10m=v10m-mean(v10m, na.rm=TRUE), 
         # mu.u10=mu.u10-mean(mu.u10, na.rm=TRUE), mu.v10=mu.v10-mean(mu.v10, na.rm=TRUE)
         ) %>% 
  mutate(Year.mod = Year-mean(Year), Year.idx = (Year-min(Year)+1))
         # u10m=mean(u10m, na.rm=TRUE), v10m=mean(v10m, na.rm=TRUE))

# Scale covariates by sd
dat = dat %>% 
  mutate(elev = elev/sd(elev), dem=dem/sd(dem), 
         slope=slope/sd(slope), 
         # u10m=u10m/sd(u10m), v10m=v10m/sd(v10m), mu.u10=sd(mu.u10), mu.v10=sd(mu.v10)
         )

# Split into training and testing sets
dat = dat %>% mutate(row.ID = 1:nrow(dat)) %>% relocate(row.ID)
dat.train = dat %>% slice_sample(prop = 0.80) %>% arrange(row.ID)
dat.test = dat %>% filter(!row.ID %in% dat.train$row.ID)
```

Directly calculate linear coefficients of time from raw data.
```{r}
yr.trends = dat.train %>% 
  group_by(trace_ID) %>% 
  do(tidy(lm(accum ~ Year.mod, data = .))) %>% 
  filter(term=='Year.mod') %>% select(-term)
dat.mu = dat.train %>% group_by(trace_ID) %>% 
  summarize(East=mean(East), North=mean(North), accum=mean(accum)) %>% 
  left_join(yr.trends %>% select(trace_ID, estimate)) %>% 
  mutate(log.est = log(1+(estimate/accum)))
```

```{r}
# More coarse mesh for development purposes
mesh = inla.mesh.2d(loc = dat.train %>% select(East, North),
                    max.edge = c(60000, 120000), cutoff = 12000)
mesh = inla.mesh.2d(loc = dat.train %>% select(East, North),
                    max.edge = c(30000, 90000), cutoff = 2000)
plot(mesh)
points(dat.train %>% select(East, North), col = "red")

# These values will also need to be adjusted when going back to Gamma?
spde = inla.spde2.pcmatern(mesh = mesh, 
  prior.range = c(20000, 0.01), # P(range < 15 km) = 0.01
  prior.sigma = c(100, 0.05))

st.idx <- inla.spde.make.index('st.idx', spde$n.spde, n.group = length(unique(dat.train$Year.idx)))
yr.idx <- inla.spde.make.index('yr.idx', spde$n.spde)

A.st <- inla.spde.make.A(mesh, 
                         loc=cbind(dat.train$East, dat.train$North), 
                         group = dat.train$Year.idx)
A.yr <- inla.spde.make.A(mesh, 
                         loc=cbind(dat.train$East, dat.train$North), 
                         weights = dat.train$Year.mod)

dat.stack <- inla.stack(
  data = list(y = dat.train$accum), 
  A = list(1, 1, A.st, A.yr),
  effects = list(list(Intercept = rep(1, nrow(dat.train))), 
                 tibble(elev = dat.train$elev, dem=dat.train$dem, 
                        # slope=dat.train$slope, aspect=dat.train$aspect, 
                        Year=dat.train$Year.mod),
                 st.idx, 
                 yr.idx),
  tag = 'dat')

# Model formula
form = y ~ -1 + Intercept + #modeling the intercept
  # dem + aspect + #fixed effects
  f(st.idx, model = spde, group=st.idx.group, 
    control.group=list(model='ar1')) + #spatial random effect with AR1 temporal autocorrelation
  f(yr.idx, model = spde) #Global effect from year and deviation from global (modeled as a spatial random effect with weights given by Year)
```

Indexing the spacetime function by year directly, rather than having two seperate terms for it (I'm not sure this will work because we don't have unique year values for each spacetime point).
```{r}
A.st <- inla.spde.make.A(mesh, 
                         loc=cbind(dat.train$East, dat.train$North), 
                         group = dat.train$Year.idx, 
                         weights = dat.train$Year.mod)

dat.stack <- inla.stack(
  data = list(y = dat.train$accum), 
  A = list(1, 1, A.st),
  effects = list(list(Intercept = rep(1, nrow(dat.train))), 
                 tibble(elev = dat.train$elev, dem=dat.train$dem, 
                        # slope=dat.train$slope, aspect=dat.train$aspect, 
                        Year=dat.train$Year.mod),
                 st.idx),
  tag = 'dat')

form = y ~ -1 + Intercept + #modeling the intercept
  f(st.idx, model = spde, group=st.idx.group, 
    control.group=list(model='ar1'))
```

## Gamma modeling

```{r mod-gamma, cache=TRUE}
# For testing, initially use gaussian

mod.gamma = inla(form,
               data = inla.stack.data(dat.stack),
               family = 'Gamma',
               control.predictor = list(compute = TRUE, A = inla.stack.A(dat.stack), link=1),
               # control.inla = list(strategy = 'simplified.laplace', int.strategy='ccd'), # Options improve computation time at cost of accuracy/precision
               control.inla = list(strategy = 'gaussian', int.strategy='eb'), 
               control.compute = list(waic=TRUE, config=TRUE))

# mod.gamma = readRDS(here('data/interim_results/mod-gamma.rds'))
summary(mod.gamma)
```

### Comparisons to GLM models

```{r}
f.glm = y ~ -1 + Intercept + Year

glm.gauss = glm(f.glm, data = dat.train %>% rename(y=accum) %>% mutate(Intercept=1, Year=Year.mod))
glm.logGaus = mod.glm = glm(f.glm, data = dat.train %>% rename(y=accum) %>% mutate(Intercept=1, Year=Year.mod), 
               family = gaussian(link = "log"))
glm.gamma = glm(f.glm, data = dat.train %>% rename(y=accum) %>% mutate(Intercept=1, Year=Year.mod), 
               family = Gamma(link = "log"))
summary(glm.gamma)
```

### Posterior checks

```{r}
# Posterior draws for year coefficients
draws=100
n.locs = spde$n.spde
mod.samples = inla.posterior.sample(draws, result = mod.gamma)
tmp = inla.posterior.sample.eval(c("yr.idx"), mod.samples)
coeff.ppc = tibble(Sample=rep(1:draws, each=n.locs), coeff=as.vector(tmp))

ggplot(coeff.ppc %>% group_by(Sample)) + 
  geom_line(aes(x=coeff, group=Sample), stat="density", alpha=0.1) + 
  geom_density(data=dat.mu, aes(x=log.est), color='red') + 
  xlim(c(-0.03, .03))
```

### Spatial comparisons

```{r}
tmp = tibble(Source="model", 
             East=mesh$loc[,1], North=mesh$loc[,2], 
             med=mod.gamma$summary.random$yr.idx$`0.5quant`, 
             LB=mod.gamma$summary.random$yr.idx$`0.025quant`, 
             UB=mod.gamma$summary.random$yr.idx$`0.975quant`) %>% 
  # mutate(weak=(LB<0 && UB>0)) %>% 
  st_as_sf(coords=c("East", "North"), crs=3031)

trends.sf = dat.mu %>% select(-trace_ID, -accum, -estimate)  %>% 
  rename(med=log.est) %>% mutate(LB=NA, UB=NA, Source="data") %>% 
  st_as_sf(coords=c("East", "North"), crs=3031) %>%
  bind_rows(tmp)

ggplot(trends.sf %>% filter(Source=="data")) + 
  geom_sf(aes(color=med)) + scale_color_gradient2(limits=c(-0.03, 0.03))
ggplot(trends.sf %>% filter(Source=="model")) + 
  geom_sf(aes(color=LB)) + scale_color_gradient2(limits=c(-0.03, 0.03))
ggplot(trends.sf %>% filter(Source=="model")) + 
  geom_sf(aes(color=med)) + scale_color_gradient2(limits=c(-0.03, 0.03))
ggplot(trends.sf %>% filter(Source=="model")) + 
  geom_sf(aes(color=UB)) + scale_color_gradient2(limits=c(-0.03, 0.03))
```
## Spacetime random effect removed

It's possible that I am removing temporal signal from my estimates when I model the spacetime auto-correlations.
Here I run the model without this random effect for comparison (so only using a spatially varying temporal coefficient and a separate temporal random effect).

```{r}
# Make data stack
stack.sep <- inla.stack(
  data = list(y = dat.train$accum), 
  A = list(1, 1, 1, A.yr),
  effects = list(list(Intercept = rep(1, nrow(dat.train))), 
                 tibble(elev = dat.train$elev, dem=dat.train$dem, 
                        slope=dat.train$slope, aspect=dat.train$aspect, 
                        Year=dat.train$Year),
                 list(time = dat.train$Year),
                 yr.idx),
  tag = 'dat')

# Priors for temporal autocorrelation
time.spec = list(rho = list(prior = 'pc.cor1', param = c(0.3, 0.95)))

# Model formula
f.sep = y ~ -1 + Intercept + 
  # f(st.idx, model = spde, group=st.idx.group, 
  #   control.group=list(model='ar1')) + 
  f(yr.idx, model = spde) + 
  f(time, model = 'ar1', hyper = time.spec) #Temporal random effect modeled as AR1

mod.sep = inla(f.sep,
                data = inla.stack.data(stack.sep),
                family = 'Gamma',
                control.predictor = list(compute = TRUE, A = inla.stack.A(stack.sep), link=1),
                control.inla = list(int.strategy='eb'), # Option improve computation time at cost of accuracy/precision
                control.compute = list(waic=TRUE, config=TRUE))

summary(mod.sep)
```

### Posterior checks

Extract marginal distributions for subset of coefficients (for plotting purposes)
```{r}
n=50
idx.set = as.integer(seq(1,spde$n.spde, length.out=n))
marg.set = mod.sep$marginals.random$yr.idx[idx.set]

# Create tbl of marginals
marg.tbl = tibble()
for (i in 1:length(idx.set)) {
  set.i = marg.set[[i]]
  tbl.i = tibble(Index = idx.set[i], X=set.i[,1], Y=set.i[,2])
  marg.tbl = marg.tbl %>% bind_rows(tbl.i)
}

ggplot(marg.tbl) + geom_line(aes(x=X, y=Y, group=Index), color='black', alpha=0.1)
```

Calculate probability that trend is less than zero.
```{r}
P.neg = vector(mode = "numeric", length = length(mod.sep$marginals.random$yr.idx))
for (i in 1:length(P.neg)) {
  marg.i = inla.tmarginal(function(x) x, 
                          mod.sep$marginals.random$yr.idx[[i]], 
                          n=500)
  P.i = diff(marg.i[,1]) * marg.i[1:(nrow(marg.i)-1),2]
  P.neg[i] = sum(P.i[which(marg.i[1:(nrow(marg.i)-1),1]<0)])
}
```

Map of estimated trends with probability of less than zero
```{r}
tmp = tibble(Source="model", 
             East=mesh$loc[,1], North=mesh$loc[,2], 
             med=mod.sep$summary.random$yr.idx$`0.5quant`, 
             P.neg=P.neg, 
             LB=mod.sep$summary.random$yr.idx$`0.025quant`, 
             UB=mod.sep$summary.random$yr.idx$`0.975quant`) %>% 
  # mutate(weak=(LB<0 && UB>0)) %>% 
  st_as_sf(coords=c("East", "North"), crs=3031)

trends.sep = dat.mu %>% select(-trace_ID, -accum, -estimate)  %>% 
  rename(med=log.est) %>% mutate(LB=NA, UB=NA, P.neg=NA, Source="data") %>% 
  st_as_sf(coords=c("East", "North"), crs=3031) %>%
  bind_rows(tmp)

ggplot(trends.sep %>% filter(Source=="data")) + 
  geom_sf(aes(color=med)) + scale_color_gradient2(limits=c(-0.05, 0.05))
# ggplot(trends.sep %>% filter(Source=="model")) + 
#   geom_sf(aes(color=LB)) + scale_color_gradient2(limits=c(-0.05, 0.05))
ggplot(trends.sep %>% filter(Source=="model")) + 
  geom_sf(aes(color=med)) + scale_color_gradient2(limits=c(-0.05, 0.05))
# ggplot(trends.sep %>% filter(Source=="model")) + 
#   geom_sf(aes(color=UB)) + scale_color_gradient2(limits=c(-0.05, 0.05))
ggplot(trends.sep %>% filter(Source=='model')) + 
  geom_sf(aes(color=P.neg)) + scale_color_viridis_c()
```

```{r}
# Posterior draws for year coefficients
draws=10
n.locs = spde$n.spde
samples.sep = inla.posterior.sample(draws, result = mod.sep)
tmp = inla.posterior.sample.eval(c("yr.idx"), samples.sep)
ppc.sep = tibble(Sample=rep(1:draws, each=n.locs), coeff=as.vector(tmp))

ggplot(ppc.sep %>% group_by(Sample)) + 
  geom_line(aes(x=coeff, group=Sample), stat="density", alpha=0.1) + 
  geom_density(data=dat.mu, aes(x=log.est), color='red') + 
  xlim(c(-0.06, .06))

```

### Spatial comparisons

## Lognormal modeling

```{r mod-gauss, cache=TRUE}

# stack.LN <- inla.stack(
#   data = list(y = log(dat.train$accum)), 
#   A = list(1, 1, A.st, A.yr),
#   effects = list(list(Intercept = rep(1, nrow(dat.train))), 
#                  tibble(elev = dat.train$elev, dem=dat.train$dem, 
#                         # slope=dat.train$slope, aspect=dat.train$aspect, 
#                         Year=dat.train$Year.mod),
#                  st.idx, 
#                  yr.idx),
#   tag = 'dat')
# 
# # For testing, initially use gaussian
# mod.gauss = inla(form, 
#                data = inla.stack.data(stack.LN), 
#                family = 'Gaussian',
#                control.predictor = list(compute = TRUE, A = inla.stack.A(stack.LN)), 
#                control.compute = list(waic=TRUE, config=TRUE))
# 
# # mod.gauss = inla(form, 
# #                  data = inla.stack.data(dat.stack), 
# #                  )
# 
# summary(mod.gauss)
```