### 2023-06-08 manuscript/processing-NB.qmd
Adds new directory for reproducible manuscript

---
title: "Processing Notebook"
format: 
  html:
    self-contained: true
editor: source
---

This notebook contains the processing workflow and instructions in order to reproduce the data, analyses, and plots included in the manuscript.

## Data pre-processing

Some pre-processing, formatting, and data wrangling are required prior to any analyses.
This includes taking the originally generated data (using PAIPR in MATLAB) and getting it into the proper format for processing in `R` using `INLA`.
Some of this processing occurs using `Python` scripts and some of it uses `R` scripts.
Both are detailed in this section.

1. Process the raw data output from PAIPR (performed in `scripts/preprocess.py`)
2. Compute topographic data necessary for later analyses (performed in `scripts/topo.calc.R`)
3. Clean the data and re-format for `R` processing (performed in `scripts/Rdata-prep.R`)
4. Build, run, and validate the `INLA` model (`scripts/inla-model.R`)
5.